---
title: "James Bond Films from Dr. No to Spectre"
subtitle: "Kaggle and Andrew Block's dataset."
author: "Sergio Aquino"
date: "01/11/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(scales)
library(janitor)
library(patchwork)
library(here)
```

## Import Data

```{r }
bond <- read_csv(here('James Bond Films','Data', 'jamesbond.csv')) %>% 
clean_names()


dim(bond) # [24,27]
glimpse(bond)
names(bond)
```

## How many films did each actor star?

```{r}

bond %>% 
  count(bond, sort = T)

```


## Film IMDB scores and actors

```{r}

bond %>% 
  group_by(bond) %>% 
  summarize(mean=round(mean(avg_user_imdb),1),
            sd=round(sd(avg_user_imdb),1)) %>% 
  arrange(-mean)

```


## James Bond directors

```{r}

bond %>% 
  count(director, sort = T)

bond %>% 
  select(year, movie, bond, director)

```

## Composer, Writer, Cinematographer

```{r}

bond %>% 
  count(composer, sort = T)

writer <- bond %>% 
  mutate_if(is.character, 
            str_replace_all, pattern = c("& | and"), replacement = ',') %>%
  select(writer) %>% 
  separate(writer, into = c("a1", "a2", "a3","a4"), sep = ",") %>% 
  pivot_longer(
    cols = starts_with("a"),
    names_to = "column",
    values_to = "writer",
    values_drop_na = TRUE
  ) %>% 
mutate(writer = str_trim(writer)) %>% 
count(writer, sort = T) 
writer

bond %>% 
  count(cinematographer, sort = T)

```

## IMBD ratings from Dr. No to Spectre 

```{r pressure, echo=FALSE}

bond %>% 
ggplot(mapping = aes(year, avg_user_imdb)) +
geom_point(data = bond) +
geom_text(data = bond, aes(label = movie), size = 4, vjust = 1, hjust = 1, nudge_x = .8, nudge_y = .05) +
   geom_line(aes(year, avg_user_imdb)) + 
  scale_fill_manual(values = c("blue", "red")) +
  xlab("") + 
  ylab("ratings") +
  geom_vline(
    aes(xintercept = year), 
    data = bond,
    colour = "grey50", alpha = 0.5
  ) +
  labs(
  title = "James Bond Movies IMBD Ratings",
  caption = "James Bond Kaggle Dataset"
)
  

```

## Bond car manufacturer, release year and actors

```{r}

bond %>% 
  count(bond_car_mfg, sort = T)

bond %>% 
  select(year, bond, bond_car_mfg) %>% 
  arrange(-year)

```


## Depicted film location

```{r}

depicted_loc <- bond %>% 
  mutate_if(is.character, 
            str_replace_all, c("England" = "United Kingdom", "Great Britain" = "United Kingdom")) %>%
  select(depicted_film_loc) %>% 
  separate(depicted_film_loc, into = c("a1", "a2", "a3","a4","a5", "a6", "a7","a8","a9"), sep = ",") %>% 
  pivot_longer(
    cols = starts_with("a"),
    names_to = "column",
    values_to = "location",
    values_drop_na = TRUE
  ) %>% 
mutate(location = str_trim(location)) %>% 
count(location, sort = T) 
depicted_loc

```

## Bond girls nationality

```{r}

girls_nat <- bond %>% 
  select(bond_girl_nat) %>% 
  separate(bond_girl_nat, into = c("a1", "a2", "a3","a4"), sep = ",") %>% 
  pivot_longer(
    cols = starts_with("a"),
    names_to = "column",
    values_to = "nationality",
    values_drop_na = TRUE
  ) %>% 
  mutate(nationality = str_trim(nationality)) %>% 
  count(nationality, sort = T) 
girls_nat 

```

## Is the mean difference between IMDB and Rotten Tomatoes ratings equal to 0. No.

```{r}
p1 <- bond %>% 
  ggplot(aes(avg_user_imdb)) +
  geom_histogram(bins = 9)

p2 <- bond %>% 
  ggplot(aes(avg_user_rtn_tom)) +
  geom_histogram(bins = 9)

bond_longer <- bond %>%
  pivot_longer(avg_user_imdb:avg_user_rtn_tom, names_to = "ratings", values_to = "values") %>% 
  select(ratings, values) 
  
p3 <-  bond_longer %>% 
  ggplot(aes(ratings, values)) +
  geom_boxplot()

(p1 | p2 | p3)
```

```{r}
bond %>% 
  summarize(avg_imdb = mean(avg_user_imdb),
            sd_imdb = sd(avg_user_imdb),
            avg_rtn = mean(avg_user_rtn_tom),
            sd_rtn = sd(avg_user_rtn_tom))

```


```{r}

t.test(values ~ ratings, data = bond_longer, paired = TRUE)

```


## Can IMDB ratings be predicted by violence, adjusted budget, number of girls per film, and film length

No, the correlation between IMDB ratings and the above variables is low.

```{r}

# create number of bond girls per film column 
library(stringi)
bond$bond_girl_film <- sapply(stri_extract_all_regex(bond$bond_girl_nat, "([^,]+)"), function(x) length(unique(x)))

bond_numeric <- dplyr::select_if(bond,is.numeric) %>% 
  select(budget_adj,film_length,avg_user_imdb,kills_bond,kills_others, bond_girl_film)

cor(bond_numeric) 

```
