---
title: "NHANES - Gun Use in the United States"
author: "Sergio Aquino"
format: html
editor: visual
---

## Objective of the analysis

The objective of the exercise is to analyze the NHANES 2011-2012 answers related to gun usage. The variables used are: incime, marital status, and race.

This exercise is based on: Harris, Statistics With R: Solving Problems Using Real-World Data, 1e SAGE Publishing, 2021

## Install libraries

```{r}
library(tidyverse)
library(here)
library(descr)
library(tableone)
library(semTools)
library(RNHANES)
library(ggmosaic)
```

## Load data from the 2011-2012 NHANES and save it locally 

```{r}
gun.usage.nhanes <- nhanes_load_data(file_name = "AUQ_G", year = "2011-2012",demographics = TRUE)

write.csv(gun.usage.nhanes, here("Exercises","data","gun.usage.nhanes.csv"))
```

## Ever used firearms? 1-Yes,2-No

The number of respondents that have never use a gun is almost twice the number of respondents that have.

```{r}
descr::freq(gun.usage.nhanes$AUQ300, plot = FALSE)

gun.usage.nhanes %>%
  filter(AUQ300 %in% c(1,2)) %>% 
  mutate(AUQ300 = ifelse(AUQ300 == 1 ,"Yes","No")) %>% 
  ggplot(aes(AUQ300)) +
    geom_bar() +
    labs(title = "NHANES Respondents that Have Used a Gun", subtitle = "Number of participants", x = "Ever used a gun?")+
    theme_light()
```

## Income breakdown of the respondents that said yes to ever using a gun

Guns are relatively inexpensive. Why do respondents with income '$100,000 and Over' have more experience with guns? 

```{r}
descr::freq(gun.usage.nhanes$INDFMIN2, plot = FALSE)

gun.usage.nhanes %>%
  select(AUQ300, INDFMIN2, DMDMARTL, RIDRETH3) %>%
  filter(AUQ300 == 1 & !(INDFMIN2 %in% c(12, 13, 77, 99))) %>%
  mutate(
    INDFMIN2 = recode_factor(
      .x = INDFMIN2,
      `1` = '$ 0 to $ 4,999',
      `2` = '$ 5,000 to $ 9,999',
      `3` = '$10,000 to $14,999',
      `4` = '$15,000 to $19,999',
      `5` = '$20,000 to $24,999',
      `6` = '$25,000 to $34,999',
      `7` = '$35,000 to $44,999',
      `8` = '$45,000 to $54,999',
      `9` = '$55,000 to $64,999',
      `10` = '$65,000 to $74,999',
      `14` = '$75,000 to $99,999',
      `15` = '$100,000 and Over'
    )
  ) %>%
  drop_na(INDFMIN2) %>%
  ggplot(aes(INDFMIN2)) +
  geom_bar() +
  coord_flip() +
  labs(title = "Income of the Respondents that Have Used a Gun", subtitle = "Levels of Income", y = "Number of Participants", x="") +
  theme_light()

```

## Marital status of respondents that have used a gun

Why have married respondents more experience with guns?

```{r}
descr::freq(gun.usage.nhanes$DMDMARTL, plot = FALSE)

gun.usage.nhanes %>%
  select(AUQ300, INDFMIN2, DMDMARTL, RIDRETH3) %>%
  filter(AUQ300 == 1 & DMDMARTL %in% c(1:6)) %>%
   mutate(
    DMDMARTL = recode_factor(
      .x = DMDMARTL,
      `1` =	'Married',
      `2` =	'Widowed',
      `3` =	'Divorced',
      `4` =	'Separated',
      `5` =	'Never married'
    )
  ) %>%
  drop_na(DMDMARTL) %>%
  mutate(DMDMARTL = DMDMARTL %>% fct_infreq()) %>% 
  ggplot(aes(DMDMARTL)) +
  geom_bar() +
  #coord_flip() +
  labs(
    title = "Marital status of the Respondents Who Have Used a Gun",
    subtitle = "NHNES Respondents",
    y = "Number of Participants",
    x = ""
  ) +
  theme_light()
```

## Race of the respondents who have used a gun

```{r}
descr::freq(gun.usage.nhanes$RIDRETH3, plot = FALSE)

gun.usage.nhanes %>%
  select(AUQ300, INDFMIN2, DMDMARTL, RIDRETH3) %>%
  filter(RIDRETH3 %in% c(1:6)) %>%
  mutate(
    RIDRETH3 = recode_factor(
      .x = RIDRETH3,
      `1` =	'Mexican American',
      `2` =	'Other Hispanic',
      `3` =	'Non-Hispanic White',
      `4` =	'Non-Hispanic Black',
      `6` =	'Non-Hispanic Asian',
      `7` =	'Other Race - Inclmutate(DMDMARTL = DMDMARTL %>% fct_infreq()) %>% uding Multi-Racial'
    )
  ) %>%
  drop_na(RIDRETH3) %>%
  mutate(RIDRETH3 = RIDRETH3 %>% 
           fct_infreq() %>% 
           fct_rev()) %>%
  ggplot(aes(RIDRETH3)) +
  geom_bar(aes(fill = RIDRETH3)) +
  scale_fill_manual(
    values = c(
      "Non-Hispanic White" = "#7463AC",
      "Non-Hispanic Black" = "gray",
      "Mexican American" = "gray",
      "Non-Hispanic Asian" = "gray",
      "Other Hispanic" = "gray"
    ),
    guide = "none"
  ) +
  labs(
    title = "Race of the Respondents Who Have Used a Gun",
    subtitle = "NHNES Respondents",
    y = "Number of Participants",
    x = ""
  ) +
  theme_light()
```

## Income and gun use

```{r}
gun.usage.nhanes %>%
  select(AUQ300, INDFMIN2, DMDMARTL, RIDRETH3) %>%
  filter(AUQ300 %in% c(1,2) & !(INDFMIN2 %in% c(12, 13, 77, 99))) %>%
  mutate(AUQ300 = ifelse(AUQ300 == 1 ,"Yes","No")) %>%
  rename(used.gun = AUQ300) %>%
  mutate(
    INDFMIN2 = recode_factor(
      .x = INDFMIN2,
      `1` = '$ 0 to $ 4,999',
      `2` = '$ 5,000 to $ 9,999',
      `3` = '$10,000 to $14,999',
      `4` = '$15,000 to $19,999',
      `5` = '$20,000 to $24,999',
      `6` = '$25,000 to $34,999',
      `7` = '$35,000 to $44,999',
      `8` = '$45,000 to $54,999',
      `9` = '$55,000 to $64,999',
      `10` = '$65,000 to $74,999',
      `14` = '$75,000 to $99,999',
      `15` = '$100,000 and Over'
    )
  ) %>%
  drop_na(INDFMIN2) %>%
  rename(income = INDFMIN2) %>%
  ggplot() +
  geom_mosaic(aes(x = product(used.gun,income), fill = used.gun)) +
  labs(title = "Income Levels and Gun Use", subtitle = "NHANES 2013", y = "", x="") +
  guides(fill=guide_legend(title="Used a gun before?")) +
  theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust=1)) +
  theme_light()


```

TBD

##

```{r}

```

