---
title: "Newspaper Stand"
editor: visual
format:
  html:
    theme: darkly
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Reading headlines

Does reading headlines from different newspapers give me a better sense of what's happening in the world than reading a few?

My goal is to spend as little time as possible navigating between newspapers. 

The most basic use of news scraping is to read headlines only. 
 
There is an assumption that you read newspapers.

## Libraries

```{r}

# install.packages("ralger") 

library(tidyverse)
library(ralger)
```

The ralger packager simplifies rvest. I used SelectorGadget to identify the html nodes I needed to grab from each newspaper page.

Deciding which newspaper headlines to read is hard.

## Wall Street Journal

```{r}
my_link <- "https://wsj.com"
my_node <- ".WSJTheme--headlineText--He1ANr9C" # The element ID , I recommend SelectorGadget if you're not familiar with CSS selectors
clean <- TRUE # Should the functionbest clean the extracted vector or not ? Default is FALSE
(wsj <- scrap(link = my_link, node = my_node, clean = clean))
```

## New York Times

```{r}
my_link <- "https://www.nytimes.com/section/todayspaper?redirect_uri=https%3A%2F%2Fwww.nytimes.com%2F"
my_node <- ".e1b0gigc0 , .e1hr934v1" 
clean <- TRUE 
(nytimes <- scrap(link = my_link, node = my_node, clean = clean))
#nytimes <- as.data.frame(nytimes)
#View(nytimes)
```

## Financial Times

```{r}
my_link <- "https://ft.com"
my_node <- ".headline--scale-3 , .stacked-secondary-story--row-stack-2 .headline--scale-2 , .text--weight-400 , .headline--scale-7" 
clean <- TRUE 
(ft <- scrap(link = my_link, node = my_node, clean = clean))
```

## Globe and Mail

```{r}
my_link <- "https://www.theglobeandmail.com/canada/"
my_node <- ".c-card__hed" 
clean <- TRUE 
(globeandmail <- scrap(link = my_link, node = my_node, clean = clean))
```

## Vancouver Sun

```{r}
my_link <- "https://vancouversun.com/"
my_node <- ".article-card__headline-clamp" # The element ID , I recommend SelectorGadget if you're not familiar with CSS selectors
clean <- TRUE # Should the functionbest clean the extracted vector or not ? Default is FALSE
(vancouversun <- scrap(link = my_link, node = my_node, clean = clean))

```

## Financial Post

```{r}
my_link <- "https://financialpost.com/category/news/"
my_node <- ".article-card__headline-clamp" 
clean <- TRUE 
(financialpost <- scrap(link = my_link, node = my_node, clean = clean))
```

Yes, reading headlines from various newspapers gives me a very good sense of what's happening around the world. It also helps me identify which news I should look into further when I open a newspaper website. 

"Early in life I had noticed that no event is ever correctly reported in a newspaper."
George Orwell

